---
title: "Data analysis and graphics for: The costs and benefits of publicising species
  discoveries."
author: "GE Ryan"
output:
  html_notebook: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r packages}
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(readxl)
library(viridis)
#library(patchwork)
library(magrittr)
library(purrr)
library(extraDistr)
library(ggforce)
library(cowplot)
```

Load functions
```{r source functions}
source("functions/calculate.zeta.R")
source("functions/cumulative.density.R")
source("functions/find.b.R")
source("functions/find.lambda.R")
source("functions/find.sigma.R")
source("functions/plot_cd_wide_zoom.R")
source("functions/plot_cd_zoom.R")
source("functions/sample.fun.R")
```


Set seed
```{r set seed}
set.seed(1984)
```


# Example scenarios

Load data
```{r load example}
example_data <- readxl::read_xlsx(
  path = "data/species_publicity_data.xlsx",
  sheet = "example_case_studies"
)

example_data
```

Calculate decision score, zeta, for each option
```{r example_results}
example_results <- example_data %>%
  as_tibble %>%
  mutate(
    zeta = benefit - cost - probability_of_leaking*(cost_of_publicity - cost),
    involve = factor(
      involve,
      levels = c(
        "Government",
        "Local community group",
        "Public"
      )
    ),
    involve = recode(
      involve,
      "Local community group" = "Local community\ngroup",
    ),
    disclose = factor(
      disclose,
      levels = c(
        "Public",
        "Part",
        "Secret"
      ),
    )
  ) %>%
  group_by(species) %>%
  mutate(
    maxval = max(zeta)
  ) %>%
  ungroup %>%
  mutate(
    optimal = ifelse(
      zeta == maxval,
      "optimal",
      "not optimal"
    )
  ) %>%
  select(-maxval)

example_results
```

Plot results from examples scenarios in figure 1.
```{r figure_1}
figure_1 <- ggplot(example_results) +
  geom_hline(
    yintercept = 0,
    colour = "darkgrey"
  ) +
  geom_point(
    aes(
      x = disclose,
      y = zeta,
      colour = involve#,
      #shape = optimal,
      #size = optimal
    ),
    size = 4#,
    # position = position_dodge(
    #   width = 0.1
    # )
  ) +
  facet_wrap(
    facets = ~ species,
    nrow = 2,
    scales = "free_y"
  ) +
  theme_grey() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_text(
      angle = 0
    ),
    axis.text.x = element_text(
      angle = 270
    )#,
    #legend.position = "bottom"
  ) +
  ylab("Decision\nscore") +
  xlab("Decision to disclose") +
  scale_colour_manual(
    values = plasma(11)[c(2,6,10)],
    guide = guide_legend(
      title = "Decision to\ninvolve",
      direction = "vertical"
    )
  ) #+
  # scale_shape_manual(
  #   values = c(19, 17),
  #   guide = guide_legend(
  #     title = "Decision is\noptimal?",
  #     direction = "vertical"
  #   )
  # ) +
  # scale_size_manual(
  #   values = c(4, 5),
  #   guide = guide_legend(
  #     title = "Decision is\noptimal?",
  #     direction = "vertical"
  #   )
  # ) 

figure_1
```

```{r figure_1 png}
png(
  filename = "plots/figure_1.png",
  width = 20,
  height = 12.5,
  units = "cm",
  res = 300
)
figure_1
dev.off()
```


```{r table_s2}
table_s2 <- example_results %>%
  select(species, disclose, involve, zeta) %>% 
  pivot_wider(
    names_from = disclose,
    values_from = zeta
  ) %>%
  arrange(species, involve)

table_s2
```

```{r write_table_s2}
write.csv(
  x = table_s2,
  file = "output/table_s2.csv",
  row.names = FALSE
)
```

# Questionnaire numeric estimates of cost and benefits
Read in numeric questionnaire data
```{r numeric_data}
numeric_data <- readxl::read_xlsx(
  path = "data/species_publicity_data.xlsx",
  sheet = "questionnaire_numeric"
)


numeric_data[ , which(substr(colnames(numeric_data), 1,1) == "q")] <- apply(
  X = numeric_data[ , which(substr(colnames(numeric_data), 1,1) == "q")],
  MARGIN = 2,
  FUN = as.numeric
)

numeric_data
```

Transform to long format and tidy
```{r numdat}
numdat <- numeric_data %>%
  select(
    -status,
    -currency,
    -consent_sp_id,
    -consent_sp_loc,
    -sp_loc
  ) %>%
  pivot_longer(
    cols = starts_with("q"),
    names_to = "question",
  ) %>%
  mutate(
    decision = substr(
      x = question,
      start = 2,
      stop = 2
    ) %>%
      as.integer,
    type = substr(
      x = question,
      start = 4,
      stop = 4
    ) %>%
      as.integer,
    estimate = substr(
      x = question,
      start = 6,
      stop = 6
    ) %>%
      as.integer
  ) %>%
  mutate(
    decision = case_when(
      decision == 1 ~ "Public",
      decision == 2 ~ "Part",
      decision == 3 ~ "Secret"
    ),
    decision = factor(
      x = decision,
      levels = c(
        "Public",
        "Part",
        "Secret"
      )
    ),
    type = case_when(
      type == 1 ~ "Benefit",
      type == 2 ~ "Cost",
      type == 3 ~ "pd"
    ),
    estimate = case_when(
      estimate == 1 ~ "lower",
      estimate == 2 ~ "upper",
      estimate == 3 ~ "most.likely",
      estimate == 4 ~ "pconf"
    )
  ) %>%
  select(
    species,
    decision,
    type,
    estimate,
    value
  ) %>%
  filter(
    species != "Species2"
  )

numdat
```

Separate out cost and benefit data, widen format, and determine type of distribution to be fit to estimates
```{r cost_benefit}
cost_benefit <- numdat %>%
  filter(type != "pd") %>%
  pivot_wider(
    names_from = estimate,
    values_from = value
  ) %>%
  mutate(
    pconf = pconf/100
  ) %>%
  filter(species != "Species2") %>%
  mutate(
    most.likely = ifelse(
      test = is.na(most.likely),
      yes = (upper - lower)/2,
      no = most.likely
    ),
    pconf = ifelse(
      test = is.na(pconf),
      yes = 0.95,
      no = pconf
    ),
    dist = case_when(
      pconf == 1 ~ "triangular",
      lower == most.likely & lower != 0 & pconf > 0.5 ~ "triangular",
      lower == most.likely ~ "exponential",
      TRUE ~ "lognormal"
    )
  )


cost_benefit
```

Estimate parameters for lognormal distributions
```{r cb_ln}
cost_benefit_lognormal <- cost_benefit %>%
  filter(dist == "lognormal") %>%
  mutate(
    value = pmap(
      .l = list(
        lower = lower,
        most.likely = most.likely,
        upper = upper,
        p = pconf
      ),
      .f = find.sigma,
      print.sigma = FALSE
    )
  ) %>%
  unnest(value)

cost_benefit_lognormal
```

Estimate parameters for exponential distributions
```{r cb_ex}
cost_benefit_exponential <- cost_benefit %>%
  filter(dist == "exponential") %>%
  mutate(
    value = pmap(
      .l = list(
        lower = lower,
        upper = upper,
        p = pconf
      ),
      .f = find.lambda,
    )
  ) %>%
  unnest(value)

cost_benefit_exponential
```

Estimate parameters for triangular distributions
```{r cb_tr}
cost_benefit_triangular <- cost_benefit %>%
  filter(dist == "triangular") %>%
  mutate(
    value = pmap(
      .l = list(
        lower = lower,
        most.likely = most.likely,
        upper = upper,
        p = pconf
      ),
      .f = find.b,
    )
  ) %>%
  unnest(value)

cost_benefit_triangular
```

Number of random samples to take from distributions
```{r nsamples}
nsamples <- 1000
```

Bind cost benefit data data back together and sample distributions for each variable
```{r cb_all}
cost_benefit_all <- bind_rows(
  cost_benefit_lognormal,
  cost_benefit_exponential,
  cost_benefit_triangular
) %>%
   mutate(
    samples = pmap(
      .l = list(
        dist = dist,
        most.likely = most.likely,
        value = value
      ),
      .f = sample.fun,
      n = nsamples
    )
  ) 

cost_benefit_all
```

Table s3. Generate table of statistical distributions and parameters used for sampling
```{r table_s3}
distribution_table <- cost_benefit_all %>%
  select(
    species,
    decision,
    type,
    lower,
    most.likely,
    everything(),
    -samples
  ) %>%
  mutate(
    sigma = ifelse(
      dist == "lognormal",
      value,
      NA
    ),
    mu = ifelse(
      dist == "lognormal",
      log(most.likely) + sigma^2,
      NA
    ),
    lambda = ifelse(
      dist == "exponential",
      value,
      NA
    ),
    a = ifelse(
      dist == "triangular",
      0,
      NA
    ),
    b = ifelse(
      dist == "triangular",
      value,
      NA
    ),
    c = ifelse(
      dist == "triangular",
      most.likely,
      NA
    )
  ) %>%
  select(
    -value
  ) %>%
  arrange(species, decision, type)

distribution_table
```

Write out table s3
```{r write_table_s3}
write.csv(
  x = distribution_table,
  file = "output/table_s3.csv",
  row.names = FALSE
)
```

Probability information becomes public given a decision, pd.
```{r pd}
pd_all <- numdat %>%
  filter(type == "pd") %>%
  mutate(
    pd = value/100
  ) %>%
  select(
    species,
    decision,
    pd
  )

pd_all
```

Cost if a decision is public, Cp
```{r cp}
cost_public <- cost_benefit_all %>%
  filter(decision == "Public") %>%
  filter(type == "Cost") %>%
  select(
    species,
    Cost_public = samples
  )

cost_public
```

Calculate estimates of decision score zeta from samples
```{r zeta_estimates}
zeta_estimates <- cost_benefit_all %>%
  pivot_wider(
    id_cols = c(
      species, decision
    ),
    names_from = type,
    values_from = samples
  ) %>%
  arrange(
    species, decision
  ) %>%
  full_join(
    cost_public
  ) %>%
  full_join(
    pd_all
  ) %>%
  mutate(
    pd = ifelse(
      test = is.na(pd),
      yes = 1,
      pd
    )
  ) %>%
  mutate(
    id = list(1:nsamples)
  ) %>%
  unnest(
    cols = c(
      Benefit,
      Cost,
      Cost_public,
      id
    )
  ) %>%
  mutate(
    zeta = pmap(
      .l = list(
        Bd = Benefit,
        Cd = Cost,
        pd = pd,
        Cp = Cost_public
      ),
      .f = calculate.zeta
    )
  ) %>%
  unnest(zeta) 

zeta_estimates
```

Minimum and maximum decision scores
```{r zeta_estimates_minmax}
zeta_estimates_minmax <- zeta_estimates %>%
  group_by(species) %>%
  summarise(
    min = min(zeta),
    max = max(zeta)
  )

zeta_estimates_minmax
```

Sequential values from min to max along each range
```{r zeta_ranges}
zeta_ranges <- zeta_estimates_minmax %$%
  mapply(
    FUN = seq,
    from = min,
    to = max,
    MoreArgs = list(
      length.out = 1000
    )
  )

colnames(zeta_ranges) <- zeta_estimates_minmax$species

zeta_ranges <- zeta_ranges %>%
  as_tibble %>%
  pivot_longer(
    cols = everything(),
    names_to = "species",
    values_to = "value"
  ) %>%
  arrange(species, value)

zeta_ranges
```

Clculate the cumlative density for each decision scenario at each value for each species
```{r zeta_cd_wide}
zeta_cd_wide <- zeta_ranges %>%
  rowwise %>%
  mutate(
    Public = cumulative.density(
      species,
      value,
      data = zeta_estimates,
      nsamples = nsamples,
      decision = "Public"
    ),
    Part = cumulative.density(
      species,
      value,
      data = zeta_estimates,
      nsamples = nsamples,
      decision = "Part"
    ),
    Secret = cumulative.density(
      species,
      value,
      data = zeta_estimates,
      nsamples = nsamples,
      decision = "Secret"
    )
  ) 

zeta_cd_wide
```

```{r zeta_dominance}
zeta_dominance <- zeta_cd_wide %>%
  group_by(species) %>%
  summarise(
    Public_Part   = sum(Public <= Part)   == nsamples,
    Public_Secret = sum(Public <= Secret) == nsamples,
    Part_Public   = sum(Part <= Public)   == nsamples,
    Part_Secret   = sum(Part <= Secret)   == nsamples,
    Secret_Public = sum(Secret <= Public) == nsamples,
    Secret_Part   = sum(Secret <= Part)   == nsamples
  ) %>%
  pivot_longer(
    cols = -species,
    names_to = "dominance",
    values_to = "dominates"
  ) %>%
  mutate(
    dominant = sub(
      pattern = "_.*",
      "",
      dominance
    ),
    dominated = sub(
      pattern = ".*_",
      "",
      dominance
    )
  ) %>%
  select(species, dominant, dominates, dominated) %>%
  pivot_wider(
    id_cols = c(species, dominant),
    names_from = dominated,
    values_from = dominates
  )

zeta_dominance
```

```{r write zeta_dominace}
# write.csv(
#   x = zeta_dominance,
#   file = "output/table_s6.csv",
#   row.names = FALSE
# )
```


```{r zeta_cd_limits}
Zeta_cd_limits <- zeta_estimates %>%
  group_by(species, decision) %>%
  summarise(
    low = quantile(
      x = zeta,
      probs = 0.1
    ),
    high = quantile(
      x = zeta,
      probs = 0.9
    )
  ) %>%
  group_by(species) %>%
  summarise(
    lower = min(low),
    upper = max(high)
  )

Zeta_cd_limits
```

Transform cumulative densitiy to long format for plotting
```{r zeta_cd}
zeta_cd <- zeta_cd_wide %>%
  pivot_longer(
    cols = c(Public, Part, Secret),
    names_to = "decision",
    values_to = "cd"
  ) %>%
  mutate(
    decision = factor(
      decision,
      levels = c(
        "Public",
        "Part",
        "Secret"
      )
    )
  )

zeta_cd
```


## Figure 2

```{r fig2_list}
fig2_list <- mapply(
  FUN = plot_cd_zoom,
  species = unique(zeta_estimates$species),
   xlab = c(
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE
  ),
  ylab = c(
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE
  ),
  # xtext = c(
  #   FALSE,
  #   FALSE,
  #   FALSE,
  #   FALSE,
  #   FALSE,
  #   TRUE,
  #   TRUE
  # ),
  xtext = TRUE,
  MoreArgs = list(
    data = zeta_cd,
    limits = Zeta_cd_limits,
    textsize = 10
  ),
  SIMPLIFY = FALSE
)

fig2_list[[1]]
```

```{r figure_2}
#f2legend <- get_legend(fig2_list[[1]])

figure_2 <- plot_grid(
  NULL,
  NULL,
  NULL,
  NULL,
  fig2_list[[1]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  fig2_list[[2]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  fig2_list[[3]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  fig2_list[[4]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  fig2_list[[5]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  fig2_list[[6]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  fig2_list[[7]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  #fdlegend,
  ncol = 4,
  align = "hv",
  labels = c(
    "",
    "",
    "",
    "",
    "Alpine tree-frog",
    "Giant Ibis",
    "Red Handfish",
    "Smooth Newt",
    "Species1",
    "Tiger",
    "Wollemi Pine"#,
    #""
  ),
  #label_fontface = "plain",
  label_size = 12,
  vjust = -0.01,
  hjust = 0,
  rel_heights = c(0.1, 1, 1)
)

figure_2
```

```{r print figure_2}
png(
  filename = "plots/figure_2.png",
  width = 20,
  height = 12.5,
  units = "cm",
  res = 300
)
figure_2
dev.off()
```

### Figure s1

```{r figs1_list}
figs1_list <- mapply(
  FUN = plot_cd_wide_zoom,
  species = unique(zeta_estimates$species),
   xlab = c(
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE
  ),
  ylab = c(
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE,
    FALSE
  ),
  # xtext = c(
  #   FALSE,
  #   FALSE,
  #   FALSE,
  #   FALSE,
  #   FALSE,
  #   TRUE,
  #   TRUE
  # ),
  xtext = TRUE,
  MoreArgs = list(
    data = zeta_cd,
    limits = Zeta_cd_limits,
    textsize = 10
  ),
  SIMPLIFY = FALSE
)

figs1_list[[1]]
```

```{r figure_s1}
fs1legend <- get_legend(figs1_list[[1]])

figure_s1 <- plot_grid(
  NULL,
  NULL,
  NULL,
  NULL,
  figs1_list[[1]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  figs1_list[[2]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  figs1_list[[3]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  figs1_list[[4]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  figs1_list[[5]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  figs1_list[[6]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  figs1_list[[7]] + scale_colour_manual(values = viridis(3)[3:1], guide = FALSE),
  #fdlegend,
  ncol = 4,
  align = "hv",
  labels = c(
    "",
    "",
    "",
    "",
    "Alpine tree-frog",
    "Giant Ibis",
    "Red Handfish",
    "Smooth Newt",
    "Species1",
    "Tiger",
    "Wollemi Pine"#,
    #""
  ),
  #label_fontface = "plain",
  label_size = 12,
  vjust = -0.01,
  hjust = 0,
  rel_heights = c(0.1, 1, 1)
)

figure_s1
```

Print figure_s1
```{r print figure_s1}
png(
  filename = "plots/figure_s1.png",
  width = 20,
  height = 20,
  units = "cm",
  res = 300
)
figure_s1
dev.off()
```


Find best option for each simulation, i.e., the one that maximises the decision score, and find whether the best option, as well as each disclosure decision would result in a value above zero, i.e., and expected net benefit.
```{r best_option}
best_option <- zeta_estimates %>%
  pivot_wider(
    id_cols = c(
      species,
      id
    ),
    names_from = decision,
    values_from = zeta
  ) %>%
  select(
    species, id, Public, Part, Secret
  ) %>%
  rowwise %>%
  mutate(
    zetamax = max(Public, Part, Secret),
    best = case_when(
      zetamax == Public ~ "Public",
      zetamax == Part ~ "Part",
      zetamax == Secret ~ "Secret"
    ),
    zetamin = min(Public, Part, Secret),
    worst = case_when(
      zetamin == Public ~ "Public",
      zetamin == Part ~ "Part",
      zetamin == Secret ~ "Secret"
    )
  ) %>%
  ungroup %>%
  mutate(
    max_positive = ifelse(
      test = zetamax > 1,
      yes = 1,
      no = 0
    ),
    pub_positive = ifelse(
      test = Public > 1,
      yes = 1,
      no = 0
    ),
    par_positive = ifelse(
      test = Part > 1,
      yes = 1,
      no = 0
    ),
    sec_positive = ifelse(
      test = Secret > 1,
      yes = 1,
      no = 0
    )
  )
  

best_option
```

Summarise data into proportions for plotting.
```{r best_option_proportion}
best_option_proportion <- best_option %>%
  group_by(species, best) %>%
  summarise(n = n()) %>%
  mutate(
    prop.best = n/1000,
    best = factor(
      x = best,
      levels = c(
        "Public",
        "Part",
        "Secret"
      )
    )
  )

best_option_proportion
```


```{r worst_option_proportion}
worst_option_proportion <- best_option %>%
  group_by(species, worst) %>%
  summarise(n = n()) %>%
  mutate(
    prop.worst = n/1000,
    worst = factor(
      x = worst,
      levels = c(
        "Public",
        "Part",
        "Secret"
      )
    )
  )

worst_option_proportion
```


## Figure 3.
Best (a) ...
```{r fig3a}
figure_3a <- ggplot(best_option_proportion) +
  geom_bar(
    aes(
      x = species,
      y = prop.best,
      fill = best
    ),
    position = "stack",
    stat = "identity"
  ) +
  ylab("Proportion of times decision is best") +
  xlab("Species") +
  scale_fill_manual(
    values = viridis(3)[3:1],
    guide = guide_legend(
      title = "Decision\nto disclose",
      direction = "vertical"
    )
    #guide = FALSE
  ) +
  ylim(c(0,1)) +
  theme(
    axis.text.x = element_text(
      angle = 270
    )
  )

figure_3a
```

... and worst (b) options. 
```{r fig3b}
figure_3b <- ggplot(worst_option_proportion) +
  geom_bar(
    aes(
      x = species,
      y = prop.worst,
      fill = worst
    ),
    position = "stack",
    stat = "identity"
  ) +
  ylab("Proportion of times decision is worst") +
  xlab("Species") +
  scale_fill_manual(
    values = viridis(3)[3:1],
    guide = guide_legend(
      title = "Decision\nto disclose",
      direction = "vertical"
    )
    #guide = FALSE
  ) +
  ylim(c(0,1)) +
  theme(
    axis.text.x = element_text(
      angle = 270
    )
    
  )

figure_3b
```

Figure 3
```{r figure_3}
figure_3 <- figure_3a +
  figure_3b  +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "a")

figure_3
```

```{r print figure_3}
png(
  filename = "plots/figure_3.png",
  width = 20,
  height = 12.5,
  units = "cm",
  res = 300
)
figure_3
dev.off()
```

Table of proportion of times best option
```{r best_option_table}
best_option_table <-  best_option_proportion %>%
  pivot_wider(
    id_cols = species,
    names_from = best,
    values_from = prop.best
  ) %>%
  select(
    species,
    Public,
    Part,
    Secret
  )

best_option_table
```

And worst option
```{r worst_option_table}
worst_option_table <-  worst_option_proportion %>%
  pivot_wider(
    id_cols = species,
    names_from = worst,
    values_from = prop.worst
  ) %>%
  select(
    species,
    Public,
    Part,
    Secret
  )

worst_option_table
```

Write out table s4a
```{r write_table_s4a}
write.csv(
  x = best_option_table,
  file = "output/table_s4a.csv",
  row.names = FALSE
)
```

Write out table s4b
```{r write_table_s4b}
write.csv(
  x = worst_option_table,
  file = "output/table_s4b.csv",
  row.names = FALSE
)
```

Table of proportion of times decision score is >0
```{r net_benefit_table}
net_benefit_table <- best_option %>%
  group_by(
    species
  ) %>%
  summarise(
    Best = sum(max_positive)/1000,
    Public = sum(pub_positive)/1000,
    Part = sum(par_positive)/1000,
    Secret = sum(sec_positive)/1000
  )

net_benefit_table
```

Write out table s5
```{r write_table_s5}
write.csv(
  x = net_benefit_table,
  file = "output/table_s5.csv",
  row.names = FALSE
)
```

Transform for plotting
```{r net_benefit_plot}
net_benefit_plot <- net_benefit_table %>%
  pivot_longer(
    cols = -species,
    names_to = "Decision",
    values_to = "Proportion positive"
  ) %>%
  mutate(
    Decision = factor(
      x = Decision,
      levels = c(
        "Best",
        "Public",
        "Part",
        "Secret"
      )
    )
  )

net_benefit_plot
```


### Figure s2
```{r figs2}
figure_s2 <- ggplot(net_benefit_plot) +
  geom_bar(
    aes(
      x = Decision,
      y = `Proportion positive`,
      fill = Decision
    ),
    stat = "identity"
  ) + 
  facet_wrap(
    facets = ~ species,
    nrow = 2
  ) +
  ylab("Proportion of times decision results in net benefit") +
  xlab("Decision to disclose") +
  scale_fill_manual(
    values = c("darkgrey", viridis(3)[3:1]),
    guide = guide_legend(
      title = "Decision\nto disclose",
      direction = "vertical"
    ) 
  ) +
  theme(
    axis.text.x = element_text(
      angle = 270
    ),
    legend.position = c(0.85, 0.2)
  ) +
  ylim(c(0,1))

figure_s2
```


```{r save figs2}
png(
  filename = "plots/figure_s2.png",
  width = 20,
  height = 12.5,
  units = "cm",
  res = 300
)
figure_s2
dev.off()
```





# Questionnaire written responses on types of costs and benefits
```{r text_dat}
text_dat <- readxl::read_xlsx(
  path = "data/species_publicity_data.xlsx",
  sheet = "questionnaire_text"
)
```

```{r text_results}
q_numeric <- sapply(text_dat, is.numeric)
sum_resp <- sapply(
  X = text_dat,
  FUN = function(x){
    
    isnum <- is.numeric(x)
    
    if(isnum){
      result <- sum(x, na.rm = TRUE)
    } else {
      result <- NA
    }
    
  }
)

perc_resp <- sapply(
  X = text_dat,
  FUN = function(x){
    
    isnum <- is.numeric(x)
    
    lenx <- length(x)
    
    if(isnum){
      result <- 100*sum(x, na.rm = TRUE)/lenx
    } else {
      result <- NA
    }
    
  }
)

text_results <- tibble(
  question = colnames(text_dat),
  sum_responses = sum_resp,
  percentage_responses = perc_resp
)

text_results
```

```{r all_results}
all_results <- text_results %>%
  filter(substr(question, 1, 3) == "all") %>%
  mutate(
    c1 = sub(
      pattern = ".{0,4}_",
      replacement = "",
      question
    ),
    c2 = sub(
      pattern = ".{0,8}_",
      replacement = "",
      c1
    ),
    cost_benefit= sub(
      pattern = "_.*",
      replacement = "",
      c1
    ),
    group = sub(
      pattern = "_.*",
      replacement = "",
      c2
    ),
    response = sub(
      pattern = ".{0,8}_",
      replacement = "",
      c2
    )
  ) %>%
  select(cost_benefit, group, response, sum_responses, percentage_responses) %>%
  mutate(
    response = recode(
      response,
      "funding" = "Funding",
      "awareness-publicsupport" = "Awareness &\npublic support",
      "protection-recovery" = "Protection &\nrecovery",
      "species" = "Species",
      "individuals" = "Individual\npeople",
      "researchersandinstitutes" = "Researchers &\ninstitutes",
      "gov" = "Government",
      "business" = "Business",
      "ngo" = "NGOs",
      "genpublic" = "General\npublic",
      "genpub" = "General\npublic",
      "poaching" = "Poaching",
      "extinctionextirpation" = "Extinction &\nextirpation",
      "disease" = "Disease",
      "funding-financial" = "Funding",
      "reputation" = "Reputation"
    )
  )

all_results
```

Split each kind for plotting
```{r separate}
benefits <- all_results %>% filter(cost_benefit == "benefits")
costs <- all_results %>% filter(cost_benefit == "costs")

benefits_kind     <- benefits %>% filter(group == "kind")
benefits_entities <- benefits %>% filter(group == "entities")
benefits_source   <- benefits %>% filter(group == "source")

costs_kind     <- costs %>% filter(group == "kind")
costs_entities <- costs %>% filter(group == "entities")
```


```{r fig4a}
figure_4a <- ggplot(benefits_kind) +
  geom_bar(
    aes(
      x = response,
      y = percentage_responses
    ),
    fill = viridis(9)[8],
    stat = "identity",
    width = 0.9*3/7
  ) +
  #ylab("Percentage of informants") +
  xlab("Kinds of benefit") +
  ylim(0, 100) +
  coord_flip() +
  theme(
    axis.title.x = element_blank()
  )

figure_4a
```

```{r fig4c}
figure_4c <- ggplot(benefits_entities) +
  geom_bar(
    aes(
      x = response,
      y = percentage_responses
    ),
    fill = viridis(9)[8],
    stat = "identity",
    width = 0.9
  ) +
  #ylab("Percentage of informants") +
  xlab("Entities\nrecieving benefits") +
  ylim(0, 100) +
  theme(
    axis.text.y = element_text(
      size = 7,
      lineheight = 0.7
    ),
    axis.title.x = element_blank()
  )+
  coord_flip()

figure_4c
```

```{r fig4e}
figure_4e <- ggplot(benefits_source) +
  geom_bar(
    aes(
      x = response,
      y = percentage_responses
    ),
    fill = viridis(9)[8],
    stat = "identity",
    width = 0.9*3/7
  ) +
  ylab("Percentage of informants") +
  xlab("Source of benefits") +
  ylim(0, 100) +
  coord_flip()

figure_4e
```

```{r fig4b}
figure_4b <- ggplot(costs_kind) +
  geom_bar(
    aes(
      x = response,
      y = percentage_responses
    ),
    fill = viridis(9)[2],
    stat = "identity",
    width = 0.9*5/7
  ) +
  ylab("Percentage of informants") +
  xlab("Kinds of cost") +
  ylim(0, 100) +
  theme(
    axis.text.y = element_text(
      lineheight = 0.7
    ),
    axis.title.x = element_blank()
  ) +
  coord_flip()

figure_4b
```

```{r fig4d}
figure_4d <- ggplot(costs_entities) +
  geom_bar(
    aes(
      x = response,
      y = percentage_responses
    ),
    fill = viridis(9)[2],
    stat = "identity",
    width = 0.9*3/7
  ) +
  ylab("Percentage of informants") +
  xlab("Entities\nbearing costs") +
  ylim(0, 100) +
  coord_flip()

figure_4d
```

```{r fig4}
figure_4 <- figure_4a +
  figure_4b +
  figure_4c +
  figure_4d +
  figure_4e +
  plot_layout(
    ncol = 2
  ) +
  plot_annotation(
    tag_levels = "a"
  )

figure_4
```

```{r save fig4}
png(
  filename = "plots/figure_4.png",
  width = 20,
  height = 15,
  units = "cm",
  res = 300
)
figure_4
dev.off()
```

